```{r echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
knitr::opts_chunk$set( echo=FALSE,warning=FALSE,message=FALSE)
```


TITLE by Anfeng Yu
========================================================


```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(dplyr)
library(GGally)

```

```{r echo=FALSE, Load_the_Data}
# Load the Data
LoanData <- read.csv("prosperLoanData.csv")

```

This data set contains 113,937 loans with 81 variables on each loan, including 
loan amount, borrower rate (or interest rate), current loan status, borrower 
income, borrower employment status, borrower credit history, and the latest 
payment information.

# Univariate Plots Section


First


```{r echo=FALSE}
#The bar graph of CreditGrade's distribution.
ggplot(aes(x = CreditGrade), data = LoanData) + geom_bar() +
  xlab("Credit Grade")
```

Most customers have no credit grade. Among customers who have credit grade, most
have C credit grade.

```{r echo=FALSE}
#The bar graph of CreditGrade's distribution.
ggplot(aes(x = ProsperScore), data = LoanData) + geom_bar() +
  xlab("Prosper Score")
```

Most customers have 3 ~ 9 prosper score.And I find the prosper scores of customers
are discrete.

```{r echo=FALSE}

#Set levels of prosper score
levels(LoanData$ProsperScore) <- c(1,2,3,4,5,6,7,8,9,10,11)

```


```{r echo=FALSE}

#The bar graph of Term's distribution.
ggplot(aes(x = Term) , data = LoanData) + geom_bar() + scale_x_continuous(breaks
= c(12,36,60)) 

```

Most customers have 36 terms loan, some have 60 terms loan, very few have 12 
terms loan.



```{r echo=FALSE}
#The bar graph of IncomeRange's distribution.
ggplot(aes(IncomeRange), data = LoanData) + geom_bar() + 
xlab("Income Range") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

Most cumtomers' income range from $25,000 ~ $74,999.


```{r echo=FALSE}
#The histogram graph of LoanOriginalAmount's distribution.
ggplot(aes(LoanOriginalAmount), data = LoanData) + geom_histogram(binwidth = 1000)+
  scale_x_continuous(breaks = seq(0,35000,5000)) + 
xlab("Loan Original Amount")

```

Most cumtomers' original loan amout is under $15000, and many of customers loan 
$5000, $10000, $15000, $20000, $25000. Very few cumstomers loan more than $25000.


```{r echo=FALSE}
#The histogram graph of LoanCurrentDaysDelinquent's distribution.
ggplot(aes(x = LoanCurrentDaysDelinquent), 
       data = LoanData[LoanData$LoanCurrentDaysDelinquent > 0,]) + 
  geom_histogram(binwidth = 50)+ 
  scale_x_continuous(breaks = seq(0,2500,200)) + 
  xlab("Loan Current Days Delinquent")


```

This graph is kind like 2 normal distribution's combination. We can make the 
guess: Among the customers who have loan delinquent, there are two major kinds,
one usually has about less delinquent days(about 200 days), but the another kind
has much more days(about 2000 days).

```{r echo=FALSE}
#The histogram graph of LoanCurrentDaysDelinquent's distribution.

ggplot(aes(x = EstimatedEffectiveYield), data = LoanData) + geom_histogram() + 
  scale_x_continuous(breaks = seq(-0.2,0.3,0.05)) + 
  xlab("Estimated Effective Yield")

```

By this graph, it's obvious that most customers' EstimatedEffectiveYield is 
higher than 0.05. Very few customers' EstimatedEffectiveYield is below 0.

```{r echo=FALSE}
#The histogram graph of LoanCurrentDaysDelinquent's distribution.

ggplot(aes(x = LenderYield), data = LoanData) + geom_histogram() + 
  scale_x_continuous(breaks = seq(-0.2,0.3,0.05)) + 
  xlab("Lender Yield")

```

The Lender Yield's distribution is very similar to Estimated Effective Yield,
except there is no negative value.



```{r echo=FALSE}
#The histogram graph of OnTimeProsperPayments's distribution.

ggplot(aes(OnTimeProsperPayments), data = LoanData) + geom_histogram() + 
  scale_x_continuous(breaks = seq(0,150,10)) + xlab("On Time Prosper Payments")

```

By this graph, most customers' OnTimeProsperPayments is below 100, and mainly
distributed around 10 and 35.


LoanCurrentDaysDelinquent has a unusual distribution, because most data are 0.

```{r echo=FALSE}
#The histogram graph of OnTimeProsperPayments's distribution.

ggplot(aes(x = LoanCurrentDaysDelinquent), data = LoanData) + 
  geom_histogram(binwidth = 50)+ 
  scale_x_continuous(breaks = seq(0,2500,200)) +
  xlab("Loan Current Days Delinquent")

```

By this graph, alomost all customers' LoanCurrentDaysDelinquent is 0, for further
investigation, I choose to use data with LoanData$LoanCurrentDaysDelinquent > 0.

```{r echo=FALSE}
#The histogram graph of OnTimeProsperPayments(above 0)'s distribution.
ggplot(aes(x = LoanCurrentDaysDelinquent), data =                             LoanData[LoanData$LoanCurrentDaysDelinquent > 0,]) + 
  geom_histogram(binwidth = 50)+ 
  scale_x_continuous(breaks = seq(0,2500,200)) +
   xlab("Loan Current Days Delinquent (days above 0)")

```

By this graph,I can see that the distribution of LoanCurrentDaysDelinquent 
above 0 is like two normal distributions' combination, one with mean 200, 
another with mean about 1950.



# Univariate Analysis



### What is the structure of your dataset?

The prosperLoanData is a set of each customers' loan and personal information 
data.For each customer, this data set has his/her loan amount and rate, salary 
and job, also estimated effictive yield(earned money).


### What is/are the main feature(s) of interest in your dataset?

By my exploration, I think main feature is estimatedEffctiveYield, because this 
feature reveal by this customer how much interest we company can make.


### What other features in the dataset do you think will help support your 
###investigation into your feature(s) of interest?

First,IncomeRange, StatedMonthlyIncome and LoanOriginalAmount, these features 
help me to figure out if this person has the ability to pay his loan.
Second OnTimeProsperPayments, creditGrade can show if this customer can usually 
pay his loan.



### Did you create any new variables from existing variables in the dataset?

No.


### Of the features you investigated, were there any unusual distributions? Did 
###you perform any operations on the data to tidy, adjust, or change the form of
###the data? If so, why did you do this?




# Bivariate Plots Section


```{r echo=FALSE}
#The bar graph of LoanCurrentDaysDelinquent agianst IncomeRange distribution.
LoanData$IncomeRange <- factor(LoanData$IncomeRange)
levels(LoanData$IncomeRange) <- c("Not displayed","Not employed","$0","$1-24,999","$100,000+","$25,000-49,999","$50,000-74,999"
                                  ,"$75,000-99,999" )


ggplot(aes(y = LoanCurrentDaysDelinquent, x = IncomeRange), data = LoanData) + 
  geom_boxplot() + ylab("Loan Current Days Delinquent")+
  xlab("Income Range") + theme(axis.text.x = element_text(angle = 90, hjust = 1))
 
```

Customers with income range of ($50,000~ $74,999) have much higher mean 
LoanCurrentDaysDelinquent.

```{r echo=FALSE}
#The bar graph of LoanCurrentDaysDelinquent agianst IncomeRange distribution, with
#LoanCurrentDaysDelinquent above 0.

ggplot(aes(y = LoanCurrentDaysDelinquent, x = IncomeRange), 
       data = LoanData[LoanData$LoanCurrentDaysDelinquent > 0,]) +
  geom_boxplot() + ylab("Loan Current Days Delinquent")+
  xlab("Income Range") + theme(axis.text.x = element_text(angle = 90, hjust = 1))


```

But among data set of customers who have already delinquent loan, they have 
similar mean of LoanCurrentDaysDelinquent.



```{r echo=FALSE}

#The boxplot of EstimatedEffectiveYield agianst IncomeRange distribution.

ggplot(aes(x = factor(IncomeRange) , y = EstimatedEffectiveYield), data = LoanData) + 
  geom_boxplot(outlier.colour = "orange") + ylab("Estimated Effective Yield")+
  xlab("Income Range") + theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

It seems there is no strong relationship between IncomeRange and EstimatedEffectiveYield.




```{r echo=FALSE}
#The scatter plot of EstimatedEffectiveYield agianst OnTimeProsperPayments distribution.

ggplot(aes(x = OnTimeProsperPayments, y = EstimatedEffectiveYield), data = LoanData) +
  geom_jitter(alpha = 0.1) + ylab("Estimated Effective Yield")+
  xlab("On Time Prosper Payments")


```

By this graph, it is hard to see if there is any strong relationship between 
OnTimeProsperPayments and EstimatedEffectiveYield.

```{r echo=FALSE} 
#computing corelationship of OnTimeProsperPayments and EstimatedEffectiveYield
with(LoanData[!is.na(LoanData$OnTimeProsperPayments) &
                !is.na(LoanData$EstimatedEffectiveYield),],
     cor(OnTimeProsperPayments,EstimatedEffectiveYield))

```

By computing corelationship, we can see there is no siginificant relationship 
between EstimatedEffectiveYield and OnTimeProsperPayments.


```{r echo=FALSE}
#The scatter plot of EstimatedEffectiveYield agianst LoanOriginalAmount distribution.
ggplot(aes(x = LoanOriginalAmount, y = EstimatedEffectiveYield), data = LoanData[!is.na(LoanData$LoanOriginalAmount)&!is.na(LoanData$EstimatedEffectiveYield),]) +
  geom_jitter(alpha = 0.1) + ylab("Estimated Effective Yield")+
  xlab("Loan Original Amount")
```

We can see there may be some relationship between LoanOriginalAmount and EstimatedEffectiveYield.

```{r echo=FALSE}
#The smooth plot of EstimatedEffectiveYield agianst LoanOriginalAmount distribution.
ggplot(aes(x = LoanOriginalAmount, y = EstimatedEffectiveYield), data = LoanData[!is.na(LoanData$LoanOriginalAmount)&!is.na(LoanData$EstimatedEffectiveYield),]) +
  geom_smooth() + ylab("Estimated Effective Yield")+
  xlab("Loan Original Amount")
```

We can see, when LoanOriginalAmount is below 15000, the EstimatedEffectiveYield 
is relatively higher.




```{r echo=FALSE}
#The corelation of EstimatedEffectiveYield agianst LoanOriginalAmount distribution.
with(LoanData[!is.na(LoanData$LoanOriginalAmount ) & !is.na(LoanData$EstimatedEffectiveYield),],
     cor(LoanOriginalAmount ,EstimatedEffectiveYield))
```

The corelation between these two features are also proof of this.


```{r echo=FALSE}
#The smooth plot of EstimatedEffectiveYield agianst ProsperScore distribution.
ggplot(aes(x = factor(ProsperScore) , y = EstimatedEffectiveYield), data = LoanData[!is.na(LoanData$ProsperScore)&!is.na(LoanData$EstimatedEffectiveYield),]) + 
  geom_boxplot(outlier.colour = "orange") + ylab("Estimated Effective Yield") +
  xlab("Prosper Score") 
```

Also, there is a strong relationship between ProsperScore and EstimatedEffectiveYield.
The higher ProsperScore is, the less EstimatedEffectiveYield is.



```{r echo=FALSE}
#The corelation between EstimatedEffectiveYield and ProsperScore distribution.
with(LoanData[!is.na(LoanData$ProsperScore ) & !is.na(LoanData$EstimatedEffectiveYield),],
     cor(ProsperScore ,EstimatedEffectiveYield))
```

The corelation also proves this.


```{r echo=FALSE}
#The smooth plot of EstimatedEffectiveYield agianst LenderYield distribution.
ggplot(aes(x = LenderYield , y = EstimatedEffectiveYield), data = LoanData[!is.na(LoanData$LenderYield)&!is.na(LoanData$EstimatedEffectiveYield),]) +
  geom_smooth() + ylab("Estimated Effective Yield")+  xlab("Lender Yield")
```

There is a very strong positive ratio relationship between LenderYield and 
EstimatedEffectiveYield, however when LenderYield is higher than 0.3, 
EstimatedEffectiveYield goes downwards.

```{r echo=FALSE}
#The corelation between EstimatedEffectiveYield and LenderYield distribution.
with(LoanData[!is.na(LoanData$LenderYield ) & !is.na(LoanData$EstimatedEffectiveYield),], 
     cor(LenderYield ,EstimatedEffectiveYield))
```

The corelation also proves this.


```{r echo=FALSE}
#The smooth plot of LenderYield agianst ProsperScore distribution.
ggplot(aes(x = factor(ProsperScore) , y = LenderYield), data = LoanData[!is.na(LoanData$ProsperScore)&!is.na(LoanData$LenderYield),]) + 
  geom_boxplot(outlier.colour = "orange") + ylab("Lender Yield")+ 
  xlab("Prosper Score")
```

The relationship is even more obvious for LenderYield and ProsperScore.

```{r echo=FALSE}
# The corelation between LenderYield and ProsperScore
with(LoanData[!is.na(LoanData$LenderYield ) & !is.na(LoanData$ProsperScore),], 
     cor(LenderYield ,ProsperScore))
```

The corelation also proves this.



# Bivariate Analysis

By bivariate explorations, I find there is a strong relationship between 
EstimatedEffectiveYield with LoanOriginalAmount and ProsperScore.

### Talk about some of the relationships you observed in this part of the \ 
###investigation.How did the feature(s) of interest vary with other features in  the dataset?

For IncomeRange in $50,000 ~ $74,999, there is no data for EstimatedEffectiveYield.


### Did you observe any interesting relationships between the other features \ 
###(not the main feature(s) of interest)?
1.The relationship is even more obvious for LenderYield and ProsperScore.
2.Customers with income range of ($50,000~ $74,999) have much higher mean 
LoanCurrentDaysDelinquent.

### What was the strongest relationship you found?
There is a very strong positive ratio relationship between LenderYield and 
EstimatedEffectiveYield, however when LenderYield is higher than 0.3, 
EstimatedEffectiveYield goes downwards.

# Multivariate Plots Section


```{r echo=FALSE, Multivariate_Plots}
#The smooth plot of EstimatedEffectiveYield agianst LenderYield distribution 
#facet by ProsperScore.
ggplot(aes(x = LenderYield, y = EstimatedEffectiveYield), data = LoanData) + 
  geom_smooth() + facet_wrap(~LoanData$ProsperScore) + 
  ylab("Estimated Effective Yield")+ xlab("Lender Yield") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))


```


These plots(x is LenderYield, y is EstimatedEffectiveYield, faceted by ProsperScore)
clearly show that, generally LenderYield has a positive ratio relationship with
EstimatedEffectiveYield, and  EstimatedEffectiveYield will decrease a little with 
ProsperScore's increasing.

```{r echo=FALSE}
#The smooth plot of EstimatedEffectiveYield agianst LoanCurrentDaysDelinquent 
#distribution facet by ProsperScore.

ggplot(aes(x = LoanCurrentDaysDelinquent, y = EstimatedEffectiveYield), data = LoanData) + 
  geom_smooth() + facet_wrap(~LoanData$ProsperScore) + 
  ylab("Estimated Effective Yield")+ xlab("Loan Current Days Delinquent") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

By this graph, it is obvious that, with higher Loan Current Days Delinquent, the
customers' Estimated Effective Yield will siginificantly drop down. But, if this
customer has high ProsperScore, the Estimated Effective Yield will change very 
little with Loan Current Days Delinquent.It can be illstrated as with higher 
ProsperScore, even if this customer has loan delinquent, bank can still estimate
a high  Effective Yield.


```{r echo=FALSE}
#The smooth plot of EstimatedEffectiveYield agianst LoanOriginalAmount distribution
#facet by ProsperScore.

ggplot(aes(x = LoanOriginalAmount, y = EstimatedEffectiveYield), data = LoanData) +
  geom_smooth() + facet_wrap(~LoanData$ProsperScore) + 
  ylab("Estimated Effective Yield")+ xlab("Loan Original Amount") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

By this graph, we can see, with higher loan original amount, the estimated effective
yield is relatively low, and with higher prosper score, the estimated effective 
yield will become lower.



# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \ 
###investigation. Were there features that strengthened each other in terms of \
###looking at your feature(s) of interest?

These plots(x is LenderYield, y is EstimatedEffectiveYield, faceted by ProsperScore)
clearly show that, generally LenderYield has a positive ratio relationship with
EstimatedEffectiveYield, and  EstimatedEffectiveYield will decrease a little with
ProsperScore's increasing.

### Were there any interesting or surprising interactions between features?

For ProsperScore between 1~7, at last EstimatedEffectiveYield will decrease with
LenderYield, but for higher ProsperScore, it is always positive ratio.

### OPTIONAL: Did you create any models with your dataset? Discuss the \ 
###strengths and limitations of your model.

------

# Final Plots and Summary



### Plot One
```{r echo=FALSE, Plot_One}
#The boxplot plot of EstimatedEffectiveYield agianst ProsperScore distribution.
ggplot(aes(x = factor(ProsperScore) , y = EstimatedEffectiveYield), data =
         LoanData[!is.na(LoanData$ProsperScore)&
        !is.na(LoanData$EstimatedEffectiveYield),])+
  geom_boxplot(outlier.colour = "orange") + 
  ylab("Estimated Effective Yield")+ xlab("Prosper Score") +
  ggtitle("boxplot plot of EstimatedEffectiveYield vs ProsperScore")

```


### Description One

there is a strong relationship between ProsperScore and EstimatedEffectiveYield.
The higher ProsperScore is, the less EstimatedEffectiveYield is.

### Plot Two
```{r echo=FALSE, Plot_Two}
#The smooth plot of EstimatedEffectiveYield agianst LenderYield distribution.
ggplot(aes(x = LenderYield , y = EstimatedEffectiveYield), data = LoanData[!is.na(LoanData$LenderYield)&!is.na(LoanData$EstimatedEffectiveYield),])+ 
  geom_smooth() + ylab("Estimated Effective Yield")+ xlab("Lender Yield") +
  ggtitle("smooth line plot of EstimatedEffectiveYield vs LenderYield")
```

### Description Two

There is a very strong positive ratio relationship between LenderYield and 
EstimatedEffectiveYield, however when LenderYield is higher than 0.3, 
EstimatedEffectiveYield goes downwards.


### Plot Three
```{r echo=FALSE, Plot_Three}
#The smooth plot of EstimatedEffectiveYield agianst LenderYield distribution 
#facet by ProsperScore.
ggplot(aes(x = LenderYield, y = EstimatedEffectiveYield), data = LoanData) + 
  geom_smooth() + facet_wrap(~LoanData$ProsperScore) + 
  ylab("Estimated Effective Yield")+ xlab("Lender Yield") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ggtitle("smooth line plot of EstimatedEffectiveYield vs LenderYield facet by ProsperScore.")
```


### Description Three

These plots(x is LenderYield, y is EstimatedEffectiveYield, faceted by ProsperScore)
clearly show that, generally LenderYield has a positive ratio relationship with
EstimatedEffectiveYield, and  EstimatedEffectiveYield will decrease a little with
ProsperScore's increasing.


------

# Reflection


For this project, my most siginificant obstacle is to find what is the main feature.
Finally I decided EstimatedEffectiveYield as my main feature because it is related
with profit the bank can earn for each customer. However, there still are some
other features very important, for example, Loan Current Days Delinquent, which
may show if this customer can pay his/her credit or not.

So, for future's further investigation, I would like to explore other features 
from other perspective, and at last make a fine predictive model to decide should
the bank  loan money to some customer or not.



